{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-06 thesis/intro.tex
minor intro edits

%Exponential explosion is easily understood through the tale of Sissa Ibn D\`{a}hir
The Islamic historian Ibn Khallik\={a}n told perhaps the first cautionary tale of exponential explosion
\cite{ibn-khallikan,khallikān1868ibn}.
Sissa Ibn D\`{a}hir,
credited with the invention of chaturanga (the precursor to modern chess),
was invited by King Shihr\^{a}m to request any reward he desired.
Sissa requested that a grain of wheat be placed on the first square of a chessboard,
two in the second, and so on, doubling the number of grains in each previous square until all 64 be filled.
The king at first laughed, thinking it a pittance,
realizing only later that to fulfill this would require
more than all the wheat in all the cities on Earth.
Modern programmers will recognize this sum as $2^{64}-1$,
the largest value representable by an unsigned 64-bit integer,
or approximately 18 billion billion.
The number of interleavings in a concurrent system grows at a similarly intractable rate
(although not necessarily exactly doubling each time),
with the number of chessboard squares filled with grain corresponding to the number of events in a program's execution.
A concurrency bug, then, is a single poisoned grain of wheat,
which must be rooted out from the stock before feeding the townspeople.
And Sissa himself is the programmer,
whose software is invariably large enough to be far beyond the reach of any exhaustive verification strategy.
}

@@

\begin{itemize}
	\item {\bf Background:} \Cref{chap:background} will present the requisite background material on concurrent programming, stateless model checking, and the various types of programs targeted by Landslide.
	\item {\bf Landslide:} \Cref{chap:landslide} explains the design and implementation of Landslide
		%the core of my stateless model checking framework,
		and all the special features it's been equipped with over the years.
		It is the foundation upon which all three contributions above build.
	\item {\bf Quicksand:} \Cref{chap:quicksand} presents the Iterative Deepening framework which more intelligently chooses which state spaces to test, corresponding to contribution 1 above.
	\item {\bf Education:} \Cref{chap:education} discusses my evaluation of Landslide
		in CMU's 15-410 class environment using the Pebbles kernel,
		and in the University of Chicago's and Berkeley's OS class environments using the Pintos kernel,
		corresponding to contribution 2 above.
	%\item {\bf Pebbles:} \Cref{chap:410} discusses my evaluation of Landslide in CMU's 15-410 class environment using the Pebbles kernel, corresponding to contribution 2 above.
	%\item {\bf Pintos:} \Cref{chap:410} discusses my evaluation of Landslide in the University of Chicago's and Berkeley's OS class environments using the Pintos kernel, corresponding to contribution 2 above.
	\item {\bf Transactional Memory:} \Cref{chap:tm} presents my extension of Landslide's concurrency model to handle transactional concurrency and the evaluation thereof, corresponding to contribution 3 above.
	\item {\bf Related Work:} \Cref{chap:relatedwork} honors my neighbours and ancestors in research spirit.
	\item {\bf Future Work:} \Cref{chap:warpzone} lists stateless model checking's remaining shortcomings and how new research might address them.
	\item {\bf Conclusion:} \Cref{chap:conclusion} provides some thoughts on the future of the field.
\end{itemize}

>>>>

%Exponential explosion is easily understood through the tale of Sissa Ibn D\`{a}hir
The Islamic historian Ibn Khallik\={a}n told perhaps the first cautionary tale of exponential explosion
\cite{ibn-khallikan,khallikān1868ibn}.
Sissa Ibn D\`{a}hir,
credited with the invention of chaturanga (the precursor to modern chess),
was invited by King Shihr\^{a}m to request any reward he desired.
Sissa requested that a grain of wheat be placed on the first square of a chessboard,
two in the second, and so on, doubling the number of grains in each previous square until all 64 be filled.
The king at first laughed, thinking it a pittance,
realizing only later that to fulfill this would require
more than all the wheat in all the cities on Earth.
Modern programmers will recognize this sum as $2^{64}$-$1$,
the largest value representable by an unsigned 64-bit integer,
or approximately 18 billion billion.
The number of interleavings in a concurrent system grows at a similarly intractable rate
(although not necessarily exactly doubling each time),
with the number of chessboard squares filled with grain corresponding to the number of events in a program's execution.
A concurrency bug, then, is a single poisoned grain of wheat,
which must be rooted out from the stock before feeding the townspeople.
And Sissa himself is the programmer,
whose software is invariably large enough to be far beyond the reach of any exhaustive verification strategy.
}

@@

\begin{itemize}
	\item {\bf Background:} \Cref{chap:background} will present the requisite background material on concurrent programming, stateless model checking, and the various types of programs targeted by Landslide.
	\item {\bf Landslide:} \Cref{chap:landslide} explains the design and implementation of Landslide
		%the core of my stateless model checking framework,
		and all the special features it's been equipped with over the years.
		It is the foundation upon which all three contributions above build.
	\item {\bf Quicksand:} \Cref{chap:quicksand} presents the Iterative Deepening framework which more intelligently chooses which state spaces to test, corresponding to contribution 1 above.
	\item {\bf Education:} \Cref{chap:education} discusses my evaluation of Landslide
		in CMU's 15-410 class environment using the Pebbles kernel,
		and in the University of Chicago's and Berkeley's OS class environments using the Pintos kernel,
		corresponding to contribution 2 above.
	%\item {\bf Pebbles:} \Cref{chap:410} discusses my evaluation of Landslide in CMU's 15-410 class environment using the Pebbles kernel, corresponding to contribution 2 above.
	%\item {\bf Pintos:} \Cref{chap:410} discusses my evaluation of Landslide in the University of Chicago's and Berkeley's OS class environments using the Pintos kernel, corresponding to contribution 2 above.
	\item {\bf Transactions:} \Cref{chap:tm} presents my extension of Landslide's concurrency model to handle transactional concurrency and the evaluation thereof, corresponding to contribution 3 above.
	\item {\bf Related Work:} \Cref{chap:relatedwork} honors my neighbours and ancestors in research spirit.
	\item {\bf Future Work:} \Cref{chap:warpzone} lists stateless model checking's remaining shortcomings and how new research might address them.
	\item {\bf Conclusion:} \Cref{chap:conclusion} provides some thoughts on the future of the field.
\end{itemize}
